Understanding Admixture Fractions

### 2015-08-11 paper/response.tex

\documentclass[11pt]{amsart}
\usepackage{graphicx,amssymb,amsthm,setspace,fullpage,natbib}
\onehalfspacing

\title{Random Union of Gametes}
\author{Mason Liang}
\date{\today}

\begin{document}
\textit{Responses in italics}
\section{Reviewer 1}
The authors present a model for understanding how moments of admixture proportions in a 
sample of individuals from a population change over time. Unlike earlier methods that allow 
inference of admixture dates from ancestry segment lengths, their method should be applicable in 
situations where local ancestry estimates cannot be reliably obtained. 
In contrast to Verdu and Rosenberg 2011, who focus on the admixture fraction for a random 
individual at one locus, Liang and Nielsen study the mean admixture fraction of a haploid 
individual’s genome, which makes their results both more intuitive to me, but also directly 
interpretable in comparison to sample population data. [Specifically, I find their definition of H 
quite meaningful, whereas I could never figure out how $H_{1,g}$ in the earlier work could take on 
values different from 0, 1/2, or 1, since it describes the admixture proportions of one individual at 
exactly one locus, but that may be my own inflexible understanding.]\\
\textit{The random variable $H_{1,g}$ only takes on values from the set $\{0, 1/2, 1\}$, but its expectation can be equal to other values in the range $[0,1]$}\\
But furthermore they incorporate drift and the random recombination 
Aside from the numerous typos, in both text and equations, I have observed no substantial 
problems with the modeling, and find the applications to African American data well-chosen. It 
would have been neat to see the method applied to an older admixture event for which local 
ancestry estimates are not possible, but I understand it would be difficult to validate such an 
inference and older admixed datasets might not be readily available. Overall, I think this is an 
interesting and relevant work.\\ 
\textit{We used African American data because the }
\\
line 57 - delete second “model” \textit{This has been deleted.}\\
lines 116-117 equations: should the $M$ rather be $G$ in the equations? \textit{Yes, we have changed them to G's}\\
line 118 - ``factions" $\rightarrow$ ``fractions" \textit{We have added the r.}\\
line 120 - ``to the left" \textit{We have added the missing ``the''}\\
line 159 - is the location of the sites in genetic distance? \textit{, this clarification has been added to the manuscript.}\\
line 166 - delete second comma at end of equation \textit{We have deleted it.}\\
line 175 - ``interference" \textit{We have corrected the spelling.}\\
line 181 - ``this assumption" could you clarify which assumption (no crossover interference?) \textit{Yes, The line now reads ``However, none of the mathematical results of this paper will require this assumption of no crossover interference.''}\\
line 210 - I would have appreciated some explanation or definition of $\kappa_n$. \textbf{TBD}\\
line 229 - ``these are similar" \textit{We have added the missing ``are''.}\\
line 249 - ``becuase" \textit{We have corrected this.}\\
line 262 - do you define which population you infer to admix into the other? I assume you infer the migration rate of Europeans into an African population, but this is not clear from the text.
\textit{We have clarified the model ``We used a 3-parameter model of constant admixture in which the admixed population is entirely of African ancestry before generation $g_start$. For $g_{start}\leq g\leq g_{stop}$, the gene-flow probability from the European source population is $s_g=s$, with $s_g=0$ otherwise. ''}\\ 
line 263 - Did you explain which values you used in eqn 1? Ie, I assume you needed to choose $L$? Or if not, could you explain? \textit{We used $R=33$. This has been included in the text.}\\
line 291 - ``European and African" \text{This has been fixed.}\\
lines 293 and 295 – ``g\_stop" $\rightarrow$ ``$g_{stop}$" \textit{These have been fixed.}\\ 
line 294 - It is possible that these individuals may choose not to self-identify as African American, so an alternate explanation is that individuals with more recent admixture may not be in this self-identified population. \\
\textit{This is an explanation we had not considered. We have added it. ``Another possible explanation for this estimate is that individuals with more recent admixture tend to not self-identify, or were otherwise not included in the ASW population. The effects of this sampling bias would be the same as the decline in gene flow we observed.''}\\
Figure 2 - I would like this figure even more if it had estimates of $k_1$ and $k_2$ for each 
simulated population.\\
\textit{Figure 2 has been updated to include the sample variance for each replicate. We did not include the sample mean for simplicity and because it is more easily estimated by eye.}\\
Figure 3 - ``but the two are very similar"\\
\textit{We have added the missing ``re''.}

\section{Review 2}
This article proposes theoretical derivations for moments of the distribution of ancestry proportions in admixed populations. This is an interesting topic, and the authors propose a general recursive method to calculate high-order moments of the distribution that was not presented before. The manuscript takes as a starting point the model of Verdu and Rosenberg (VR) for the estimation of variance in admixed populations. VR does not take into account drift or the finiteness of the genome, and the present manuscript sets out to model these effects. The authors also distinguish between expectations within population (i.e., given a fixed population genealogy), and between populations (i.e., between individuals having independently sampled genealogies), and propose an inference scheme for structure data.

Even though this is an interesting manuscript, a few issues should be addressed.\\
\subsection{}
Gravel (2012) also discusses mean and variance in ancestry proportion in models with finite genome, drift,
and continuous migrations, although not all three jointly (pp 611-613 and 618-619). It also presents an
application to admixture timing in the ASW, also finding that variance in admixture proportion supports
more than a single pulse of admixture. The method was also used in Botigue et al 10.1073/pnas.1306223110.
This manuscript is more detailed than Gravel (2012) on this topic, but there is considerable overlap. At
the very least, the authors must acknowledge the overlap and compare the results in some detail.\\
\subsection{}
The equation at line 176.5 is comparable to the second term in equation (8) in Gravel (2012).
However, the first term of equation (8) (dominant in many practical cases) appears to be missing. Even
though the present manuscript takes into account correlations among individuals induced by the
genealogy, it does not seem to take into account the correlation among loci induced by genealogies.
If I understand things correctly, the authors start off with a "Diploid Wright-Fisher" model, but then
perform calculations according to a "Markovian Wright-Fisher" model (see Figure 2B in Gravel (2012)). If I
*don't* understand things correctly, I suppose that clarifications about exactly what expectations are
calculated would be useful (see below).

Fortunately, even if I am correct, I believe that the discrepancy can be corrected in the present models
using the law of total variance. At line 245.5, the authors do something similar, but condition the variance
over "population" genealogies. They could solve most of this problem by conditioning over "individual"
genealogies, as in Gravel (2012).

\subsection{}
The manuscript presents a moments method to estimate parameters directly from the distribution of
inferred admixture proportions. Even though distributions of ancestry proportions are indeed informative
of demography, its higher moments are also very sensitive to a diversity of effects that are not accounted
for here, such as assortative mating, population structure, nonrandom sampling, sex-biased admixture,
and correlated errors in ancestry assignments. If this method is to be considered for practical inference,
these caveats should be presented clearly to avoid mis-use. I have found practical application of variance
estimates to be perilous in most practical cases.

\subsection{}
By contrast, I suspect that drift and within-vs-cross-population effects will be small to negligibleexcept
for very small populations. This is not to say that it is not useful to model these, but it may be useful to
point this out. The authors show in figure 2 important differences in the mean ancestry proportion after
500 generations in an admixed population, but I found little detail as to how the simulation was
conducted. Gravel (2012) estimates the standard deviation in ancestry proportion across replicate
populations at 0.05 for a population of size 100, so I'm assuming that the simulation presented here uses
a population of similarly small size.

\subsection{}
Finally, if simulations have been performed, it would be a good idea to provide comparisons with the
theoretical results!

\subsection*{Minor points/clarifications}\ \\
Line 116.5: M is not defined. It may also be useful to spell out "the expectations are taken over ..."--I'm not
sure I understand exactly what was done. My understanding is that M is the population genealogy (which
yields a given distribution G), and that in the LHS terms have expectations taken over both populations and
individuals, whereas the RHS is taken over individuals in a fixed population.\\
Line 134-136
It could be good to clarify which expectations are used here as well, and what terms are being estimated.
For example, E[k1] corresponds to the LHS term at line 116.5 . However, E[k2] is an estimator for the RHS
term, since we are taking a sample mean and not the expected mean. It may be good to be explicit.\\
Line 144.5
It'd be useful to point the reader to the appendix for the definitions of the v's.\\
Line 249 "because"\\
\\
I don't always sign my reviews, but I gave so many hints at this point that I might as well--\\
Simon Gravel

\section{Reviewer 3}
Liang and Nielsen consider the joint distribution of admixture proportions $H_{i(g)}$, $i\in 1,\ldots, n$ of
n individuals sampled from an admixed population g generations after the start of admixture. Previous
work apparently explored the marginal distribution of the admixture proportion of a single individual
drawn from an admixed population. (This distribution was marginal both in the sense of applying,
marginally, to a single individual, and also being marginalized over all possible genealogies). By contrast,
Liang and Nielsen explore the joint distribution of H's conditional on the unknown genealogy connecting
the samples.

They derive expressions for the expected values of various k-statistics that can be computed from a
sample of H's, and then, by relating these expected values to admixture parameters in some special cases
they develop a method of moments estimator for these admixture parameters.

This is an interesting effort to relate admixture proportions, which are routinely estimated by programs
like structure (they are the individual q values), to the parameters of mechanistic models of population
admixture and could provide a valuable tool for increasing the interpretability of results from structure.

Below I provide a few comments that I hope will be helpful.

\subsection{} I found the author's implicit invocation of de Finetti's representation theorem and its extensions by
Hamos and Savage a little hard to wrap my head around here, primarily because it is not quite clear to me
what they are considering $\mathcal{G}$ to be. At lines 104 and 105 they seem to be making the case that
conditional on a certain realization of a random distribution the H's will be iid. But, then, later, they seem
to be equating $\mathcal{G}$ with a realization from the distribution of possible genealogies. I guess the
confusing part for me here is that if one is trying to compute the expected value of $k_2$, then presumably
you are conditioning on a certain realization of $\mathcal{G}$, and if that is the case and you are adhering to
the interpretation of $\mathcal{G}$ as being a realization from a random distribution a la the representation
theorem, then $H_{i(g)}$ and $H_{j(g)}$ should be conditionally independent and hence the terms involving the
expectations of their products would simplify quite a bit further than what we find above line 137. So, on
the one hand the authors are saying that the H's are conditionally iid, and on the other they are saying that
they are correlated and that is central to their derivation. It would be great if this could be clarified
somewhat.

\subsection{} One question I have as I read this is, "How far from sufficient are the H's?" The Hs themselves are going
to be functions of the underlying "cluster of origin" process estimate by, say, structure, when using the
linkage model. So, instead of boiling everything down to a single real number between 0 and 1 for each
individual (the H's), one might consider trying to do inference from, say, the distribution of the lengths of
tracts of common descent from different clusters. I suspect that might carry considerably more
information that is relevant to estimating admixture-model parameters than just the H's. How much
information is lost by just considering the H's? Some discussion of this would be appreciated.

\subsection{} Given 2, it seems that this method would be particularly appropriate in cases where a map is not
available --- i.e. in molecular ecology contexts studying non-model organisms. This being the case, it
would be good if the authors were able to address how robust (or not) the model might be to certain
departures from the sampling model and the mechanistic admixture model they employ. Of most concern
to me as regards the application of their method in molecular ecology contexts is the effect of population
structure within the admixed population. Population structure within the admixed population will clearly
alter the observed distribution of the H's. I think it would be good to know how the strength of an effect
like population structure compares to the strength of the admixture parameters in their model in altering
the distribution of the H's. I suspect that it could have a large effect that could complicate estimation of
the admixture parameters in populations with unknown spatial gradients of introgression, etc.

\subsection{} Echoing 3, a full set of simulations evaluating the power of the method across a range of scenarios (in
terms of number of markers and the differentiation between the two admixing populations) would be nice to
have, as well as assessments of the robustness (or not) of the model to population structure within the
admixed population. In some ways, it appears to me that the estimation of these sorts of admixture
parameters could be analogous to the estimation of K---the number of clusters---using the program
structure, an exercise that is fraught with peril and warrants caution because real populations seldom
behave like idealized ones. It would be nice if the caveats involved in this method could be stated up front.
(not that this would necessarily keep molecular ecologists from abusing it somewhat...the estimation of K
being a case in point).

\subsection{} The assumption that the estimation error of $H_{i(g)}$ is independent of $H_{i(g)}$ at line 208 seems like it
might not be the best assumption. A more reasonable assumption might be that the variance of $\epsilon_i$
is proportional to $H_i(1-H_i)$, as is common when dealing with proportions. Of course, this complicates the
calculations. Perhaps some type of variance-stabilzing transformation (like arc-sine square root) might be
useful.\\
\textit{TBD}

\subsection{} It would be nice to have a github repository with the all the scripts the authors used so that their
analyses are easily reproducible. In fact, it seems like some simple software (in R) for example,
implementing the authors' method is conspicuously absent. Many journals today require that data sets be
archived to insure reproducibility, and I think it is incumbent upon theoreticians, likewise, to offer clean
software or scripts to ensure reproducibility, and in some cases ease interpretation of methods.
\textit{The simulation code and scripts to reproduce the plots in the manuscript have been uploaded to github:  \texttt{https://github.com/z2trillion/local\_ancestry\_moments}.}

\subsection*{Minor points}
\subsubsection{} Equation above line 117. Should the $\mathcal{M}$ be $\mathcal{G}?$ If not, $\mathcal{M}$ needs definition. \textit{We have changed the $\mathcal{M}$ to a $\mathcal{G}$.}
\subsubsection{} I think that the second summation in the equation below line 136 should be over $i\neq j$, rather than
$i,j = 1$ up to n.
\textit{We have made this change.}
\subsubsection{} Line 148: Wait a moment (what a terrible pun!). Isn't $k_2$ estimating a central moment? Perhaps some more general discussion of $k$-statistics earlier in the paper would be good. 
\textit{The $k$-statistics are estimators for cumulants: $\mathbb{E}(k_i) = \kappa_i$ while the $h$-statistics are estimators for central moments: $\mathbb{E}(h_i) = \mu_i$. For $i=1,2,3$, the central moment and cumulant are the same, with $h_i$ and $k_i$ having the same formula, but for $i\geq 4$, the $i^\text{th}$ cumulant is not equal to the $i^\text{th}$ central moment.}
\subsubsection{} The manuscript would benefit from a few extra rounds of proofreading.
\end{document}

### 2015-08-11 paper/manuscript.tex
Add response to reviewers and edits to manuscript

\documentclass[11pt]{amsart}
\usepackage{graphicx,amssymb,amsthm,setspace,lineno}
\usepackage[round]{natbib}
\linenumbers
\onehalfspacing
\title{Estimating Admixture Parameters from Admixture Fractions}
\author{Mason Liang and Rasmus Nielsen}
\date{\today}

@@

\section*{Introduction}
It is common in population genetic analyses to consider individuals as belonging fractionally to two or more discrete source populations. The proportion of an individual's genome that belongs to a population is called that individual's ``admixture fraction" or ``admixture proportion". Programs such as \verb|Structure| \citep{pritchard2000inference}, \verb|Eigenstrat|, \verb|Frappe| \citep{tang2005estimation}, or \verb|Admixture| \citep{alexander2009fast} can jointly estimate these admixture fractions for multiple individuals in a sample, along with the corresponding  allele frequencies in each of the source populations. These admixture fractions are often presented in a `structure plot,' an example of which is shown in Figure \ref{structure_plot}. We will henceforth refer to these methods as 'structure analyses'. 
This approach has proven highly useful for understanding genetic relationships in many different species, e.g. humans \citep{rosenberg2002genetic}, cats \citep{menotti2008patterns}, or pandas \citep{zhang2007genetic}. Other analyses reconstruct admixture tracts for each genome in the sample, by inferring the local ancestry of every position, or window, in each sampled genome \citep{tang2006reconstructing,maples2013rfmix}. In this context, the admixture fraction for a genome is the fraction of its total length that is inherited from a particular source population.

Although structure analyses are not tied to any particular mechanistic model of population history and demography, the admixture fractions and admixture tracts are commonly interpreted to be the result of past admixture events in which modern populations were formed by admixture (or introgression) between ancestral source populations. The distribution of admixture tract lengths has been related to specific mechanistic models of admixture \citep{falush2003inference,tang2006reconstructing,pool2009inference}, and has been used to estimate times of admixture \citep{gravel2012population}. However, the admixture proportions themselves also contain information regarding admixture times. Right after an admixture event, the variance in admixture proportions within a population will be high, but will thereafter decrease, and will eventually (as we will show) converge to zero in the limit of large genomes.  
The variance in admixture fractions among individuals contains substantial information about the time since admixture that can be used in addition to the tract length distribution.  
It may in some cases be more robust than inferences based on tract lengths, because the length distribution of tracts is often difficult to infer, and is often not modeled accurately by the hidden Markov model (HMM) methods used to infer tract lengths \citep{liang2014lengths}.
Even in cases where tract lengths can be accurately inferred, studies aimed at estimating admixture times should benefit from using both variance in admixture proportions among individuals and overall admixture tract lengths distributions.

@@

We extend the model model in \citet{verdu2011general} to incorporate the effects of recombination and genetic drift by adding a a random union of gametes component. Recombination is important because even if one half of a chromosome's ancestors are from the first source population, it is unlikely that exactly one half of that chromosome's genetic material is inherited from that population. Genetic drift is important because the individuals in a sample might share ancestors and, therefore, have more similar admixture fractions than expected by chance in a model without drift. The results developed in this paper should be directly applicable for quantifying the results of a structure analysis.

\section*{The General Mechanistic Model}
We start by considering admixture fractions in haploid genomes. These haploid admixture fractions can later be paired up to create diploid admixture fractions. The admixture fraction of a (haploid) genome $H_i$, is the proportion of $H_i$ that is inherited from a particular source population. For notational simplicity, we only consider gene-flow only from one population into another. We will later discuss how to extend this model to multiple admixing source populations. We use the same mechanistic admixture model of \citet{verdu2011general}, and will use its notation where possible. Finally, we use the random union of gametes model for genetic drift and recombination, and assume a sample size of $n$ chromosomes from a single population.

In the model of \citet{verdu2011general}, a hybrid population of $N$ diploid individuals forms in generation 1 from two isolated source populations. In this first generation, individuals in the hybrid population are from the first source population with probability $s_0$ or from the second source population with probability $1-s_0$. In subsequent generations, each individual in the hybrid population draws two recombined gametes at random from the previous generation. In generation $g+1$, each gamete is, independently, from the first source population with introgression probability $s_g$, or from the hybrid population with probability $1-s_g$. If the individual inherits both of its gametes from the hybrid population, the two gametes are from the same individual with probability $1/N$, and from different individuals with probability $1-1/N$.

Mathematically, if the stochastic process $A(\ell)$ represents the local ancestry along a chromosome as a function of physical position ($\ell$):

$$
A(\ell) = \left\{
     \begin{array}{l}
       0\ :\ \ell $ descended from first source population$\\
       1\ :\ \ell $ descended from second source population$\\
     \end{array}
   \right.,
$$

then the fraction of the chromosome descended from the second source population is given by

@@

Assume that $g$ generations after the start of admixture we have randomly sampled $n$ chromosomes from the hybrid population and determined their corresponding admixture fractions, $H_{1(g)},H_{2(g)},\dots,H_{n(g)}$.
We are interested in the joint distribution of these $n$ random variables.
When $n=1$ and as $L\rightarrow\infty$, this is the admixture fraction considered in \citep{verdu2011general}.

Because the $n$ chromosomes have possibly overlapping geneologies, the admixture fractions are not independent.
However, the joint distribution of the admixture fractions does not depend on their ordering, so they are exchangeable (see appendix for mathematical definition). 
As a result, they can be viewed as being identically and independently ($\textit{iid}$) drawn from a random distribution $\mathcal{G}$.
This random distribution can be interpreted as a deterministic function of the random genealogy of the entire hybrid population up to $g$ generations in the past.
When $g$ is small, the genealogies of the $n$ samples will be unlikely to differ from $n$ non-overlapping binary trees, so $\mathcal{M}$ will be approximately constant. If $g$ is large however, this will no longer hold.

@@

\begin{align*}
\mathbb{E}(H_{1(g)})&\neq\mathbb{E}(H_{1(g)}|\mathcal{M})=\lim_{n\rightarrow\infty}\frac{1}{n}\sum_{i=1}^n H_{i(g)}\\
\text{var}\left(H_{1(g)}\right)&\neq\text{var}(H_{1(g)}|\mathcal{M})=\lim_{n\rightarrow\infty}\frac{1}{n-1}\sum_{i=1}^n \left(H_{i(g)}-\frac{1}{n}\sum_{j=0}^n H_{j(g)}\right)^2.
\end{align*}

and similarly for higher-order moments.  In other words, the moments of the admixture factions have two components: randomness due sampling of population history, and randomness due to the sampling of individuals.  The expressions to left account for both, while the expressions to the right only account for the latter.  Variances among individuals within a population will correspond to var$(H_{1(g)}|\mathcal{M})$ but var$(H_{1(g)})$ will tend to be larger than the expected sample variance calculated from multiple individuals sampled from the same population, and will rarely be useful for inference purposes.

@@

We let $k_1$ be the sample mean:

@@

We let $k_2$ be the unbiased estimator of the sample variance:

@@

\begin{align*}
	\mathbb{E}(k_2)
	&=\frac{1}{n-1}\sum_{i=1}^n \mathbb{E}(H_{i,g}^2)-\frac{1}{n(n-1)}\sum_{i,j=1}^n \mathbb{E}(H_{i,g}H_{j,g})\\
	&=\mathbb{E}(H_{1,g}^2)-\mathbb{E}(H_{1,g}H_{2,g}).
\end{align*}

These expectations can be written in terms of two-point correlation functions:

@@

$$\mathbf{v}_{2(g)}=\left( \begin{array}{r}
			\mathbb{P}\left\{A_{1,g}(\ell)=1, A_{1,g}(\ell')=1\right\} \\ \mathbb{P}\left\{A_{1,g}(\ell)=1, A_{2,g}(\ell')=1\right\}
		\end{array} \right),$$

we get that

@@

In general, the $i^\text{th}$ $k$-statistic is an unbiased estimator of the $i^\text{th}$ cumulant of $\mathcal{M}$, and its expectation can be written as an integral over $[0,L]^i $ of a linear combinations of $i$-point correlation functions.
For example,

\begin{align*}
	\mathbb{E}(k_3)&=\frac{1}{L^3}\int_0^L\int_0^L\int_0^L\left( \begin{array}{ccccc}
			1 & -1 &-1 &-1 &2	\end{array} \right)\mathbf{v}_{3(g)} d\ell d\ell'd\ell''\\
	\mathbb{E}(k_4)&=\frac{1}{L^4}\int_0^L\int_0^L\int_0^L\int_0^L \Bigl(
			1 \  \underbrace{-1}_\text{4 times} \  \underbrace{-1}_\text{3 times} \  \underbrace{2}_\text{6 times} \ 6 \Bigr)\mathbf{v}_{4(g)} d\ell d\ell'd\ell''d\ell'''\\
			\dots
\end{align*}

Remarkably, the linear combinations required to compute the expectations of the $k$-statistics correspond exactly to the higher-order disequilibria as defined in [Bennett 1948?].
Furthermore, if instead the we choose to compute the expectations of the $h$-statistics, which estimate the central moments, the linear combinations would correspond to the higher-order disequilibria as defined in [Slatkin 1972].

@@

\subsection*{A Single Admixture Event}
We start with a simple case, where introgression only occurs in the founding generation, i.e. $s_g=0$ for $g>0$. Using the random union of gametes model, we can compute $\mathbf{v}_{2(g)}$ in terms of the probabilities from the previous generation:

If two sites at $\ell$ and $\ell'$ are on the same chromosome in generation $g$, then they were inherited from one gamete from one parent in the previous generation, while if they 	are on different haplotypes, then the probability that they are from gametes from one parent is $1/N$.
On one gamete, the probability that these two sites are inherited from the same parental chromosome is $[\ell \ell']$ and the probability that they are instead inherited from different parental chromosomes is $[\ell|\ell']$.
In matrix notation,

\begin{align*}
\mathbf{v}_{2(g)}&=
		\left(\mathbf{L}_2\mathbf{U}_2\right)^g
		\mathbf{v}_{2(0)},
\end{align*}

@@

\begin{align*}
	\mathbf{U}_2&=\left( \begin{array}{cc}
					 {[\ell \ell']} & [\ell|\ell']\\
					 0 & 1
				\end{array} \right)\\
	\mathbf{L}_2&=\left( \begin{array}{cc}
					 1 & 0\\
					 1/N & 1-1/N
				\end{array} \right).
\end{align*}

This is the the same matrix equation (Wright 1933 and Hill and Robertson 1966) derived for the decay of two-locus linkage disequilibrium.
The ``alleles" we consider are the local ancestry at $\ell$ and $\ell'$.
To the extent possible, our notation will follow (Hill 1974), whose results for measures of multi-locus linkage disequilibria we use.
The matrices $\mathbf{L}_2$ and $\mathbf{U}_2$ share $(1\ -1)$ as a left-eigenvector, with corresponding eigenvalues $1-1/N$ and $[\ell\ell']$.
As a result,

>>>>

\documentclass[11pt]{amsart}
\usepackage{graphicx,amssymb,amsthm,setspace,lineno}
\usepackage[round]{natbib}
\linenumbers
\onehalfspacing
\title{Understanding Admixture Fractions}
\author{Mason Liang and Rasmus Nielsen}
\date{\today}

@@

\section*{Introduction}
It is common in population genetic analyses to consider individuals as belonging fractionally to two or more discrete source populations. The proportion of an individual's genome that belongs to a population is called that individual's `admixture fraction' or `admixture proportion'. Programs such as \verb|Structure| \citep{pritchard2000inference}, \verb|Eigenstrat| \citep{price2006principal}, \verb|Frappe| \citep{tang2005estimation}, or \verb|Admixture| \citep{alexander2009fast} can jointly estimate these admixture fractions for multiple individuals in a sample, along with the corresponding  allele frequencies in each of the source populations. These admixture fractions are often presented in a `structure plot,' an example of which is shown in Figure \ref{structure_plot}. We will henceforth refer to these methods as `structure analyses'. 
This approach has proven highly useful for understanding genetic relationships in many different species, e.g. humans \citep{rosenberg2002genetic}, cats \citep{menotti2008patterns}, or pandas \citep{zhang2007genetic}. Other analyses reconstruct admixture tracts for each genome in the sample, by inferring the local ancestry of every position, or window, in each sampled genome \citep{tang2006reconstructing,maples2013rfmix}. In this context, the admixture fraction for a genome is the fraction of its total length that is inherited from a particular source population.

Although structure analyses are not tied to any particular mechanistic model of population history and demography, the admixture fractions and admixture tracts are commonly interpreted to be the result of past admixture events in which modern populations were formed by admixture (or introgression) between ancestral source populations. The distribution of admixture tract lengths has been related to specific mechanistic models of admixture \citep{falush2003inference,tang2006reconstructing,pool2009inference}, and has been used to estimate times of admixture \citep{gravel2012population}. However, the admixture proportions themselves also contain information regarding admixture times. Following an admixture event, the variance in admixture proportions within a population will be high, but will thereafter decrease, and will eventually converge to zero in the limit of large genomes.  
The variance in admixture fractions among individuals contains substantial information about the time since admixture that can be used in addition to the tract length distribution.  
In some cases, this may be more robust than inferences based on tract lengths, because the length distribution of tracts is often difficult to infer, and is often not modeled accurately by the hidden Markov model (HMM) methods used to infer tract lengths \citep{liang2014lengths}.
Even in cases where tract lengths can be accurately inferred, studies aimed at estimating admixture times should benefit from using both variance in admixture proportions among individuals and overall admixture tract lengths distributions.

@@

We extend the model in \citet{verdu2011general} to incorporate the effects of recombination and genetic drift by adding a a random union of zygotes component. Recombination is important because even if one half of a chromosome's ancestors are from the first source population, it is unlikely that exactly one half of that chromosome's genetic material is inherited from that population. Genetic drift is important because the individuals in a sample might share ancestors and, therefore, have more similar admixture fractions than expected by chance in a model without drift. The results developed in this paper should be directly applicable for quantifying the results of a structure analysis.

\section*{The General Mechanistic Model}
We start by considering admixture fractions in haploid genomes. These haploid admixture fractions can later be paired up to create diploid admixture fractions. The admixture fraction of a (haploid) genome $H_i$, is the proportion of $H_i$ that is inherited from a particular source population. For notational simplicity, we only consider gene-flow only from one population into another. We will later discuss how to extend this model to multiple admixing source populations. We use the same mechanistic admixture model of \citet{verdu2011general}, and will use its notation where possible. Finally, we use the random union of zygotes model, with a diploid population size of $N$ ($2N$ chromosomes), for genetic drift and recombination, and assume a sample size of $n$ chromosomes from a single population.

In this model, a hybrid population of $N$ diploid individuals forms in generation 1 from two previously isolated source populations. In this first generation, individuals in the hybrid population are from the first source population with probability $s_0$ or from the second source population with probability $1-s_0$. In generation $g+1$, each chromosome is, independently, from the first source population with introgression probability $s_g$, or from the hybrid population with probability $1-s_g$. Chromosomes inherited from the hybrid population are the product of the recombination of the two chromosomes of one individual (zygote), chosen uniformly at random. Finally, these $2N$ chromosomes are paired up to form the $N$ individuals in generation $g+1$.

Finally, we let the stochastic process $A(\ell)$ represent the local ancestry along a chromosome as a function of $\ell$, the physical position:

$$
A(\ell) = \left\{
     \begin{array}{l}
       0\ :\ \ell $ is descended from first source population$\\
       1\ :\ \ell $ is descended from second source population$\\
     \end{array}
   \right. .
$$

The fraction of the chromosome descended from the second source population is given by

@@

Assume that $g$ generations after the start of admixture we have randomly sampled $n$ chromosomes from the hybrid population and determined their corresponding admixture fractions, $H_{1(g)},H_{2(g)},\dots,H_{n(g)}$.
We are interested in the joint distribution of these $n$ random variables.
When $n=1$ and as $L\rightarrow\infty$, this is the admixture fraction considered by \citet{verdu2011general}.

Because the $n$ chromosomes have possibly overlapping geneologies, the admixture fractions are not independent.
However, the joint distribution of the admixture fractions does not depend on their ordering, so they are exchangeable. 
As a result, they can be viewed as being identically and independently ($\textit{iid}$) drawn from a random distribution $\mathcal{G}$.
This random distribution can be interpreted as a function of the random genealogy of the entire hybrid population up to $g$ generations in the past.
When $g$ is small, the genealogies of the $n$ samples will be unlikely to differ from $n$ non-overlapping binary trees, so $\mathcal{G}$ will be approximately constant. If $g$ is large however, these genealogies are likely to overlap, and this will no longer be true.

@@

\begin{align*}
\mathbb{E}(H_{1(g)})&\neq\mathbb{E}(H_{1(g)}|\mathcal{G})=\lim_{n\rightarrow\infty}\frac{1}{n}\sum_{i=1}^n H_{i(g)}\\
\text{var}\left(H_{1(g)}\right)&\neq\text{var}(H_{1(g)}|\mathcal{G})=\lim_{n\rightarrow\infty}\frac{1}{n-1}\sum_{i=1}^n \left(H_{i(g)}-\frac{1}{n}\sum_{j=0}^n H_{j(g)}\right)^2,
\end{align*}

and similarly for higher-order moments.  The moments of the admixture fractions have two components: randomness from sampling the population genealogy, and randomness from the sampling of chromosomes.  The expressions to the left account for both, while the expressions to the right only account for the latter. Variances among individuals within one population correspond to var$(H_{1(g)}|\mathcal{G})$, while variances over replicate populations correspond to var$(H_{1(g)})$. This latter value will be larger than the expected sample variance calculated from multiple individuals sampled from the same population, and will rarely be useful for inference purposes.

@@

Let $k_1$ be the sample mean:

@@

Similarly, let $k_2$ be the unbiased estimator of the sample variance:

@@

\begin{align*}
	\mathbb{E}(k_2)
	&=\frac{1}{n-1}\sum_{i=1}^n \mathbb{E}(H_{i,g}^2)-\frac{1}{n(n-1)}\sum_{i\neq j} \mathbb{E}(H_{i,g}H_{j,g})\\
	&=\mathbb{E}(H_{1,g}^2)-\mathbb{E}(H_{1,g}H_{2,g}).
\end{align*}

These expectations can be written in terms of two-point correlation functions of $A$:

@@

$$\mathbf{v}_{2(g)}=\left( \begin{array}{r}
			\mathbb{P}\left\{A_{1(g)}(\ell)=1, A_{1(g)}(\ell')=1\right\} \\ \mathbb{P}\left\{A_{1(g)}(\ell)=1, A_{2(g)}(\ell')=1\right\}
		\end{array} \right),$$

we find that

@@

In general, the $i^\text{th}$ $k$-statistic is an unbiased estimator of the $i^\text{th}$ cumulant of $\mathcal{G}$, and its expectation can be written as an integral over $[0,L]^i $ of a linear combinations of $i$-point correlation functions.
For example,

\begin{align*}
	\mathbb{E}(k_3)&=\frac{1}{L^3}\int_0^L\int_0^L\int_0^L\left( \begin{array}{ccccc}
			1 & -1 &-1 &-1 &2	\end{array} \right)\mathbf{v}_{3(g)} d\ell d\ell'd\ell''\\
	\mathbb{E}(k_4)&=\frac{1}{L^4}\int_{[0,L]^4}\Bigl(
			1 \  \underbrace{-1}_\text{4 times} \  \underbrace{-1}_\text{3 times} \  \underbrace{2}_\text{6 times} \ 6 \Bigr)\mathbf{v}_{4(g)} d\ell d\ell'd\ell''d\ell'''\\
			\dots
\end{align*}

Remarkably, the linear combinations required to compute the expectations of the $k$-statistics correspond exactly to the higher-order disequilibria as defined by \citet{bennett1952theory}.
Furthermore, if instead the we choose to compute the expectations of the $h$-statistics, which estimate the central moments, the linear combinations would correspond to the higher-order disequilibria as defined by \citet{slatkin1972treating}.

@@

\subsection*{A Single Admixture Event}
We start with a simple case, where introgression only occurs in the founding generation, i.e. $s_g=0$ for $g>0$. Using the random union of zygotes model, we can compute $\mathbf{v}_{2(g)}$ in terms of the probabilities from the previous generation:

If two sites at genetic distances $\ell$ and $\ell'$ are on the same chromosome in generation $g+1$, then they were inherited from one chromosome from generation $g$ with probability $[\ell\ell']$ and from two chromosomes from generation $g$ with probability $[\ell|\ell']$.
If they are on different chromosomes, then the probability that they are descended from one chromosome in generation $g$ is $\frac{1}{2N}[\ell\ell']$ and the probability that they are descended from two chromosomes is $\frac{1}{2N}[\ell|\ell']+\left(1-\frac{1}{2N}\right)$
In matrix notation,

\begin{align*}
\mathbf{v}_{2(g+1)}&=
		\left(\mathbf{L}_2\mathbf{U}_2\right)
		\mathbf{v}_{2(g)}
		= \left(\mathbf{L}_2\mathbf{U}_2\right)^g
		\mathbf{v}_{2(0)},
\end{align*}

@@

\begin{align*}
	\mathbf{L}_2&=\left( \begin{array}{cc}
					 1 & 0\\
					 \frac{1}{2N} & 1-\frac{1}{2N}
				\end{array} \right)\\
		\mathbf{U}_2&=\left( \begin{array}{cc}
					 {[\ell \ell']} & [\ell|\ell']\\
					 0 & 1
				\end{array} \right).
\end{align*}

This is the the same matrix equation (Wright 1933 and Hill and Robertson 1966) derived for the decay of two-locus linkage disequilibrium.
The `alleles' we consider are the local ancestry at $\ell$ and $\ell'$.
To the extent possible, our notation will follow (Hill 1974), whose results for measures of multi-locus linkage disequilibria we use.
The matrices $\mathbf{L}_2$ and $\mathbf{U}_2$ share $(1\ -1)$ as a left-eigenvector, with corresponding eigenvalues $1-\frac{1}{2N}$ and $[\ell\ell']$.
As a result,
